The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2014-02-15 paper/ms.tex
paper update! almost the final version

We use the full width at half maximum (FWHM) to quantify the line
broadening. Figure \ref{fig:widthvsvelocity} shows how the FWHM increases
with rotational velocity both for the central an homogeneous source
distributions.  We parameterize the dependency of the line width as
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.9 \pm 0.041$ and $\lambda_{\rm h}=
1.9\pm 0.062$ respectively.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f6.eps}
  \includegraphics[width=0.40\textwidth]{f7.eps}
\end{center}
  \caption{Effects of rotation on the escape fraction. Left
    panel. Escape fraction as a function of rotational velocity. All
    these models have $\tau_{a}=1$. The continuous (dashed) lines
    correspond to central (homogeneous) models. Right panel. Escape
    fraction as a    function of the product $(a\tau_{\rm
      H})^{1/3}\tau_{a}$. The    analytic solution for the infinite
    slab is shown as a continuous    line. Different lines correspond
    to different rotational    velocities.     \label{fig:efvsv}} 
\end{figure*}

%\begin{figure}
%  \caption{

%    {\bf IS THE ALBEDO TAKEN INTO ACCOUNT?}
%    \label{fig:efvsNeufeld}}   
%\end{figure}

@@

In the right panel of Figure \ref{fig:efvsv} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{a}$, an approximation that is valid only
in the limit $a\tau_{\rm   H}\gg 1$. The dashed lines in the right panel of Figure
\ref{fig:efvsv} show the results for the different rotational
velocities. First of all we note that the escape fraction does not
increase from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, a fact
that is explained since there is a transition from an opaque to an extremely opaque medium
which affects how the \ly photons escape from the medium: single flight 
vs. single excursion. {\bf Que quiere decir eso de single flight y
  single excursion, porque explica eso la fraccion de escape?}

@@

However, we find some cases where there is a possible signature of
three peaks. This is only present in some models ($\tau = 10^6,
V=100$\kms and $\tau = 10^7, V=300$\kms) and could be the result of
two competing effects: the rise in the line's center as the photons in
the photosphere escape in a single excursion and the broadening in the
two peaks as the photons closer from the center have to scatter to
escape. We examine in greater detail these cases in the next
sub-section.

>>>>

We use the full width at half maximum (FWHM) to quantify the line
broadening. Figure \ref{fig:widthvsvelocity} shows how the FWHM increases
with rotational velocity both for the central an homogeneous source
distributions.  We parametrize the dependency of the line width as
\begin{equation}
 {\rm FWHM}^2 = {\rm FWHM}_{0}^2 + V_{\rm max}^2/\lambda^2,
\label{eq:fwhm}
\end{equation}
%
where FWHM$_{0}$ is the
velocity width in the static case and $\lambda$ is a positive scalar
to be determined as a fit to the data. With this test we want to know
to what extent the new velocity width can be expressed as a quadratic
sum of the two relevant velocities in the problem. We fit
simultaneously all the points corresponding to central and homogeneous
models to find $\lambda_{\rm c} = 3.9 \pm 0.041$ and $\lambda_{\rm h}=
1.9\pm 0.062$ respectively.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.40\textwidth]{f6.eps}
  \includegraphics[width=0.40\textwidth]{f7.eps}
\end{center}
  \caption{Effects of rotation on the escape fraction. Left
    panel. Escape fraction as a function of rotational velocity. All
    these models have $\tau_{a}=1$. The continuous (dashed) lines
    correspond to central (homogeneous) models. Right panel. Escape
    fraction as a    function of the product $(a\tau_{\rm
      H})^{1/3}\tau_{A}$, where $\tau_{A}$ is taking into account the albedo. The    analytic solution for the infinite
    slab is shown as a continuous    line. Different lines corresponds
    to different rotational velocities for the homogeneous models.     \label{fig:efvsv}} 
\end{figure*}

@@

In the right panel of Figure \ref{fig:efvsv} we put these results in the context of
the analytic solution for the infinite slab \citep{Neufeld90}. In
Neufeld's set-up the analytic solution depends uniquely on the product
$(a\tau_{\rm   H})^{1/3}\tau_{A}$ where $\tau_{A} = (1 - A)\tau_{a}$ and $A$ is the albedo
, an approximation that is valid only
in the limit $a\tau_{\rm   H}\gg 1$. The dashed lines in the right panel of Figure
\ref{fig:efvsv} show the results for the different rotational
velocities for the homogeneous models. First of all we note that the escape fraction does not
increase significantly from $\tau_{\rm H}=10^{5}$ to $\tau_{\rm H}=10^{6}$, a fact
that is explained since there is a transition from an opaque to an extremely opaque medium
which affects how the \ly photons escape from the medium.

@@

However, we find some cases where there is a possible signature of
three peaks. This is only present in some models ($\tau_{\rm H} = 10^6,
V=100$\kms and $\tau_{\rm H} = 10^7, V=300$\kms) and could be the result of
two competing effects: the rise in the line's center as the photons in
the photosphere escape in a single excursion and the broadening in the
two peaks as the photons closer from the center have to scatter to
escape. We examine in greater detail these cases in the next
sub-section.
